---
title: 'vildehaye sensors tutorials: Calculate Altitude by Pressure'
author: "sivan margalit"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---
## credits
This code is a translation of Sivan Toledo's MATLAB code  
The data, contributed by Dr. Yoni Vortman and Anat Levi, of navigating study done in Hula valey at 2020

## Background
After loading Temperature and pressure measurements from vildehaye sensors, you can calculate altitude by pressure data. 

For this calculation we need information about the environment barometric data on same day. We can download this information from the Israely Meteorological Service (IMS)

Accessing to IMS data requires a Token (access key).  [here a link](https://ims.gov.il/he/ObservationDataAPI) for more information.

### project structure
<ol>
<li>Data Folder:
  <ul> 
  <li>binary data from battery and barometric sensors.</li>
  <li>excel file of meteorology stations, downloaded from IMS</li> </ul>
</li>
  
<li>Function folder:
R file with 3 local functions</li>
<li>VHsensors_calcAltitude.Rproj : R project file, define the folder, where it located, as a working folder for R project</li>
</ol>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(always_allow_html=TRUE)
ims.token='f058958a-d8bd-47cc-95d7-7ecf98610e47'
```
## R Code sample 
### loading libraries and local sources
The package "toolsForAtlas" contains the methods for read binary sensors data, and access to remote database by url address.

in functions folder we contains local source files contains functions, which are not in the package

```{r load libraries, message = FALSE}
library(toolsForAtlas) # useful functions for atlas users
options("digits"=14)   # show long numbers till 14 digits
source("functions/pressures.R", encoding = "UTF-8")
```

### loading data from binary files
```{r extract_data_1, message=c(1:4)}
# loading battery data
bat_df<-read_BAT("data/tag-972001006611-BATMON.bin")
bat_df$TAG<-"972001006611"
head(bat_df, 4)

# loading barometric pressure data (in mbar units)
pressure_df<-read_BME("data/tag-972001006611-BME280.bin")
pressure_df$TAG<-"972001006611"
head(pressure_df, 4)
```

### plot baromeric data
The file functions/pressures.R contains "plot_BME" function plotting barometric data.
This function has 4 arguments :
1. bme.df - is the sensor's data
2. reference.df - optional . a data frame of environment  barometric data, as a reference. when this data frame is supply the function add it to the BME plot
3. fromTime - optional. if it not supply the fromTime is the minimum time in the dataframe
4. toTime - optional. if it not supply the toTime is the maximum time in the dataframe
```{r plot_bme_1}
bme.plot.1<-plot_BME(bme.df=pressure_df)
bme.plot.1
```

### read local meteorologic data
The file functions/pressures.R contains "get.ISM.data" function which download pressure and temperature data, from a close meteorologic station.  
In this example - it uses stattion in Har-Knaan, in Tzfat 
```{r read_IMS}
# ims.token<- Insert token from IMS, and uncomment this line
ism.lst<-get.ISM.data(easting=257000,
                      northing=780000,
                      fromTime=min(pressure_df$TIME),
                      toTime=max(pressure_df$TIME),
                      ims.token)

ism.pressure.df<-ism.lst[["pressure"]]
ism.temperatures.df<-ism.lst[["temeratures"]]
hsl<-ism.lst[["hsl"]]

print(sprintf("Station Heigh relative to sea level=%d meter",hsl))
head(ism.pressure.df)
head(ism.temperatures.df)
```
(If you get red warning from red.xlsx - it is harmless)
lets plot the barometric data, including the refence data, retreived from the local IMS station
```{r plot_bme_2}
bme.plot.2<-plot_BME(bme.df=pressure_df, 
                     reference.df = ism.pressure.df)
bme.plot.2
```

### Estimate Altitude by pressure
the function sensorsPressureToAltitude() in functions/pressures.R calculate altitude, using the formula TBD
This function has 4 arguments :
<ol>
<li>pressure_df - is the sensor's data</li>
<li>ism.pressure.df - a data frame of environment  barometric data, as a reference.</li>
<li>refTemperature - temperature drymeasurements in IMS station.</li>
<li>refAltitude - IMS station high above sea level.</li>
</ol>
```{r PressureToAltitude}
alts.df<-sensorsPressureToAltitude(pressure=pressure_df,
                                   refPressure=ism.pressure.df,
                                   refTemperature=ism.temperatures.df,
                                   refAltitude=hsl)
```

now - show the altitude plot
```{r plot_altitude}
alts.df$dateTime<-as.POSIXct(alts.df$TIME, tz="UTC", origin="1970-01-01")
ggplot(alts.df, aes(x=dateTime, y=alts))+
  geom_point(aes(colour = factor(TAG)), size=0.5)+
  theme_bw()+
  theme(legend.title = element_blank())
```